Hydra Battles and Cie

### 2021-03-08 doc/chapter-primrec.tex
correction of typos

\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
\begin{Coqsrc}
Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
\end{Coqsrc}

@@

Indeed, (\texttt{evalList $m$ $v$ $f$}) is the application to the vector $v$ of
an uncurried version of $f$.

In Library\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}, we introduce a lighter notation.

@@

Since The Ackermann function is strictly monotonous in its first argument\footnote{almost always\dots}, we may replace the large inequality with a strict one.

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}.}

@@

Please note that, if the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
$\texttt{Ack}\,x\,y\leq \texttt{Ack}\,n\,(\texttt{max}\,x\,y)$ holds.
Thus, our impossibility proof is just a sequence of easy small steps.

>>>>

\noindent
\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
\begin{Coqsrc}
Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
\end{Coqsrc}

@@

Indeed, (\texttt{evalList $m$ $v$ $f$}) is the application to the vector $v$ of
an uncurried version of $f$.
In Library\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}, we introduce a lighter notation.

@@

We prove also a strict version of this lemma, thanks to the following property (proved in Library\href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}~).

@@

If the Ackermann function were primitive recursive, then there would exist some natural number $n$, such that, for all $x$ and $y$, the inequality 
$\texttt{Ack}\,x\,y\leq \texttt{Ack}\,n\,(\texttt{max}\,x\,y)$ holds.
Thus, our impossibility proof is just a sequence of easy small steps.

### 2021-03-08 doc/part-hydras.tex
renaming of a lemma
correction of typos

Hydras quite frequently contain  multiple  copies of the same pattern. The following functions
will help us to describe and reason about replications in hydra battles.

@@

Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,

@@

@@

\begin{remark}
  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists.

\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}

@@

@@

\section{The Length of Minimal Large Sequences}

@@

%\subsection{Looking for a function definition}

Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider the functions which associate to each stritly positive integer $k$ the number $l$, where 
the interval $[k,l)$ is $\alpha$-mlarge.

@@

\begin{Coqsrc}
Definition fun_le f g  := forall n:nat,  f n <=  g n.
\end{Coqsrc}

@@

Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$ (at $k$) into a step to
$\canonseq{\lambda}{k}$, then to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.

@@

Applying the previous lemmas on successors and limit ordinals, 
we get several correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.

@@

As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.

@@

\begin{Coqsrc}
Lemma Propp284 : forall alpha beta : E0, 
   beta o< alpha -> dominates (F_ alpha) (F_ beta).
\end{Coqsrc}

>>>>

Hydras quite frequently contain  multiple adjacent  copies of the same subtree. The following functions
will help us to describe and reason about replications in hydra battles.

@@

Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$s.} and there exists some strictly positive integer $s_i$
such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$.

@@

We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.

@@

\begin{remark}
  Ketonen and Solovay~\cite{KS81} consider  large finite \emph{sets} of natural numbers,  but they are mainly used as sequences. Thus, we chosed to represent them explicitely as (sorted) lists. 
\end{remark}

The following function ``gnaws'' an ordinal $\alpha$, folowing a sequence of indices (ignoring the $0$s).
\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Paths.html\#gnaw}{Epsilon0.Paths}}

@@

Minimal large sequences can be directly defined in terms of the
predicate \texttt{path\_to} (\vref{sect:path-to-def}) which already prohibits paths containing non-final \texttt{zero}s.

@@

\section{Length of Minimal Large Sequences}

@@

Let $0<\alpha<\epsilon_0$ be an ordinal term. We consider any  function which associates maps  any strictly positive integer $k$ to the number $l$, where 
the interval $[k,l)$ is $\alpha$-mlarge.

@@

\begin{Coqsrc}
Definition fun_le f g  := forall n:nat,  f n <=  g n.
Infix "<<=" := fun_le (at level 60).
\end{Coqsrc}

@@

Let $\lambda<\epsilon_0$ be any limit ordinal. In a similar way as for successors, we decompose any path from $\lambda$  into a first step to
$\canonseq{\lambda}{k}$, followed by a path to $0$. In the following section, we assume that there exists à correct function for $\canonseq{\lambda}{k}$, \emph{for any strictly positive $k$}.

@@

Applying the previous lemmas on successors and limit ordinals, 
we obtain a few  correct implementations of \texttt{(L\_spec $\alpha$)} for small values of $\alpha$.

@@

As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}, which states that the function \texttt{F\_} is ``monotonous'' w.r.t. its first argument and ``domination''.

@@

\begin{Coqsrc}
Lemma F_mono_l: forall alpha beta : E0, 
   beta o< alpha -> dominates (F_ alpha) (F_ beta).
\end{Coqsrc}
